---
title: "NYC Taxi Interactive EDA "
author: "Damiano"
date: "14082017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



Lets read some libraries
```{r include = FALSE}



library(data.table)
library(dplyr)
library(ggplot2)
library(flexdashboard)
library(TraMineR)
#install.packages("highcharter")
#install.packages("zoo")
library(highcharter)
library(DT)
library(flexdashboard)
library(leaflet)
library(rmarkdown)
library(DT)

library(treemap)
library(viridisLite)
```


Then read datasets
I decided to use dataset with weather in NYC in 2016, so it require some data wrangling to join this 2 datasets. 




```{r echo = FALSE}


train_dataset <- fread("../input/nyc-taxi-trip-duration/train.csv")
train_dataset[, pi_dt_shift :=  paste(substr(pickup_datetime, 1, 13), ":00:00", sep = "")]
train_dataset[, df_dt_shift := paste(substr(dropoff_datetime, 1, 13), ":00:00", sep = "")]


weather_nyc <-  fread("../input/knycmetars2016/KNYC_Metars.csv")



```

Lets see  how Conditions in weather_nyc dataset look like

```{r echo = FALSE}



weather_condition_freq <- weather_nyc %>%
                          group_by(Conditions) %>%
                          select(Conditions ) %>%
                          summarize(count = n())


datatable(weather_condition_freq)
```


Next step is to join 2 datasets and show some summary statistics 

```{r echo = FALSE}

train_joined <- dplyr::left_join(train_dataset, weather_nyc, by  = c("pi_dt_shift" = "Time"))

train_joined$Conditions[is.na(train_joined$Conditions) ==  TRUE] <- "Unknown"



weather_condition_freq <- train_joined %>%
  group_by(Conditions) %>%
  select(Conditions,trip_duration ) %>%
  summarize(count = n(), 
            mean_dur = mean(trip_duration, na.rm = TRUE), 
            sd_dur =   sd(trip_duration, na.rm = TRUE), 
            median_dur = median(trip_duration, na.rm = TRUE))


datatable(weather_condition_freq)

```


Plot below shows how frequently users pickup taxi depend on different weather condition ^^

I decide to shift Conditions value with NA to "Unknown category"

it is obvious that the most frequent group is group with 'Clear' Condition.

```{r}

highchart()%>%
hc_add_series(weather_condition_freq, "column", hcaes(x =  Conditions, y = count), name = "Count by Conditions Weather") %>%
  hc_plotOptions(series = list(
    showInLegend = FALSE, 
    pointFormat = "{point.y}%"
  ), 
  column = list(colorByPoint = TRUE)) %>%
  hc_subtitle(text = "Count by Conditions Caegories") %>%
  hc_credits(
    enabled = TRUE, 
    text = "Source: Kaggle", 
    href = "https://kaggle.com/damianpanek", 
    style = list(fontSize = "12px")
  ) %>%
  hc_add_theme(hc_theme_google())



```



The plot below show simple summary statitics - Mean/Median and SD of TRIP Duration


```{r echo = FALSE}



highchart()%>%
  hc_add_series(weather_condition_freq, "spline", hcaes(x =  Conditions, y = mean_dur), name = "Mean Trip Duration") %>%
  hc_add_series(weather_condition_freq, "spline", hcaes(x =  Conditions, y = median_dur), name = "Median Trip Duration") %>%
  hc_add_series(weather_condition_freq, "spline", hcaes(x =  Conditions, y = sd_dur), name = "SD Trip Duration") %>%
  hc_plotOptions(series = list(
    showInLegend = TRUE, 
    pointFormat = "{point.y}%"
  ), 
  column = list(colorByPoint = TRUE)) %>%
  hc_subtitle(text = "Count by Conditions Caegories") %>%
  hc_credits(
    enabled = TRUE, 
    text = "Source: Kaggle", 
    href = "https://kaggle.com/damianpanek", 
    style = list(fontSize = "12px")
  ) %>%
  hc_add_theme(hc_theme_google())
```


Little data transformation. I want to get information of day/month and 
day of week of observation.

```{r echo = FALSE}



train_joined <- data.table(train_joined)

train_joined <- train_joined[is.na(pickup_datetime) == FALSE,  ]




train_joined[, pickup_datetime := as.POSIXct(pickup_datetime, format = "%Y-%m-%d %H:%M:%S")]
train_joined[, dropoff_datetime := as.POSIXct(dropoff_datetime, format = "%Y-%m-%d %H:%M:%S")]
train_joined[, pickup_day := format(pickup_datetime, "%Y-%m-%d")]
train_joined[, pickup_month := format(pickup_datetime, "%Y-%m")]


train_joined[, dropoff_day := format(dropoff_datetime, "%Y-%m-%d")]
train_joined[, dropoff_month := format(dropoff_datetime, "%Y-%m")]

train_joined[, weekday := weekdays(pickup_datetime)]

```


Summary Statistics for Tempertarure in NYC   taxi dataset
```{r}



weather_temp_day <-  train_joined %>% 
  group_by(pickup_day) %>%
  select(pickup_day, Temp., Conditions) %>%
  summarize(count = n(), 
            min = min(Temp., na.rm = TRUE), 
            max = max(Temp., na.rm = TRUE), 
            sd_dur = sd(Temp., na.rm = TRUE))


datatable(weather_temp_day)


```

```{r echo = FALSE}


hchart(weather_temp_day, 
        type = "columnrange", 
        hcaes(x = pickup_day, low = min, high = max, color = sd_dur)) %>%
        hc_chart(polar = TRUE) %>%
    hc_yAxis(max = 30,  min = -10, labels = list(format = "{value} "), 
             showFirstLabel = FALSE) %>%
  hc_xAxis(
  title = list(text = ""), gridLineWidth = 0.5,
  labels = list(format = "{value: %b}")) %>%
  hc_add_theme(hc_theme_google()) %>%
hc_title(text = "Min/Max temperature daily, coloured by SD(Temp)")

```


Similar plot - Summary statistics for Trip Duration variable


```{r echo = FALSE}

weather_dur_day <-  train_joined %>% 
  group_by(pickup_day) %>%
  select(pickup_day, trip_duration, Conditions) %>%
  summarize(count = n(), 
            median = median(trip_duration, na.rm = TRUE), 
            mean = mean(trip_duration, na.rm = TRUE), 
            sd_dur = sd(trip_duration, na.rm = TRUE))


datatable(weather_dur_day)


```


```{r echo = FALSE}


hchart(weather_dur_day, 
       type = "columnrange", 
       hcaes(x = pickup_day, low = mean, high = median, color = median)) %>%
  hc_chart(polar = TRUE) %>%
  hc_yAxis( max = 1300, labels = list(format = "{value} "), 
           showFirstLabel = FALSE) %>%
  hc_xAxis(
    title = list(text = ""), gridLineWidth = 0.5,
    labels = list(format = "{value: %b}")) %>%
  hc_add_theme(hc_theme_google()) %>% 
  hc_title(text = "Trip duration Statistics per day")


```


Piechart for fwd/store flag 




```{r echo = FALSE}




store_and_fwd_freq <- train_dataset %>% 
  select(store_and_fwd_flag) %>%
  group_by(store_and_fwd_flag) %>%
  summarize(count = n()) %>%
  mutate(freq = count/sum(count))



datatable(store_and_fwd_freq)


```



```{r echo = FALSE}
hc <-  highchart() %>%
      hc_add_series(store_and_fwd_freq, "pie", hcaes(x =  store_and_fwd_flag, y = count), name = "Column Plot") %>%
  hc_plotOptions(series = list(
    showInLegend = FALSE, 
    pointFormat = "{point.y}%"
  ), 
  column = list(colorByPoint = TRUE)) %>%
  hc_subtitle(text = "Frequency of Store And FWD FLAG") %>%
  hc_credits(
    enabled = TRUE, 
    text = "Source: Kaggle", 
    href = "https://kaggle.com/damianpanek", 
    style = list(fontSize = "12px")
  ) %>%
  hc_add_theme(hc_theme_google())



```


Frequency plot - day by day 

```{r echo = FALSE}

freq_by_day <- train_joined %>%
              select(pickup_day) %>%
              group_by(pickup_day) %>%
              summarize(count = n())


datatable(freq_by_day)

```


```{r echo = FALSE}
freq_day <- highchart() %>%
            hc_add_series(freq_by_day, "column", 
                          hcaes(x = pickup_day, y = count),name = "Column")  %>%
                          hc_add_theme(hc_theme_google()) %>%
                          hc_plotOptions(
                            series = list(
                              showInLegend = FALSE, 
                              pointFormat = "{point.y}%"
                            ), 
                            column = list(
                              colorByPoint = TRUE
                            )
                          ) %>% 
  hc_yAxis(title = list("pickup per Day"), 
           labels = list(format = "{value}"))   %>%
  hc_xAxis(unique(as.character(freq_by_day$pickup_day))) %>%
  hc_title(
    text = "Graph represents amount of pickups per day"
  ) %>%
  hc_subtitle(text = "In sweet rainbow dash taste XD") %>%
  hc_credits(
    enabled = TRUE, text = "Damiano ;p/click",
    href = "https://www.kaggle.com/damianpanek"
  ) %>%
  hc_add_theme(hc_theme_google())

freq_day
```

Similar plot but observation divided by month

```{r echo =FALSE}



freq_by_month <- train_joined %>%
  select(pickup_month) %>%
  group_by(pickup_month) %>%
  summarize(count = n())


datatable(freq_by_month)
```


```{r echo  = FALSE}
freq_month <- highchart() %>%
  hc_add_series(freq_by_month, "column", 
                hcaes(x = pickup_month, y = count),name = "Column")  %>%
  hc_add_theme(hc_theme_google()) %>%
  hc_plotOptions(
    series = list(
      showInLegend = FALSE, 
      pointFormat = "{point.y}%"
    ), 
    column = list(
      colorByPoint = TRUE
    )
  ) %>% 
  hc_yAxis(title = list("pickup per Month"), 
           labels = list(format = "{value}"))   %>%
  hc_xAxis( unique(as.character(freq_by_month$pickup_month))) %>%
  hc_title(
    text = "Graph represents amount of pickups per day"
  ) %>%
  hc_subtitle(text = "UP 20170723") %>%
  hc_credits(
    enabled = TRUE, text = "Damiano ;p/click",
    href = "https://www.kaggle.com/damianpanek"
  )



freq_by_month
```



Summary statistics for trip_duration - daily


```{r echo = FALSE}



freq_by_day_trip <- train_joined %>%
  select(pickup_day, trip_duration) %>%
  
  group_by(pickup_day) %>%
  summarize(count = n(), 
            mean_trip = mean(trip_duration, na.rm = TRUE), 
            median_trip = median(trip_duration, na.rm = TRUE), 
            sd_trip     = sd(trip_duration, na.rm = TRUE))


datatable(freq_by_day_trip)

```


```{r echo = FALSE}


hc_by_day <- highchart() %>%
  hc_plotOptions(
    series = list(
      showInLegend = FALSE, 
      pointFormat = "{point.y}%"
    ), 
    column = list(
      colorByPoint = TRUE
    )
  ) %>% 
  highchart() %>%
  hc_add_series(freq_by_day_trip, "line",  hcaes(x = pickup_day, y = mean_trip),name = "Mean") %>%
  hc_add_series(freq_by_day_trip,   "line" , hcaes(x=  pickup_day,  y= median_trip), name = "median") %>%
  hc_add_series(freq_by_day_trip, "line", hcaes(x =  pickup_day, y = sd_trip), name = "sd") %>% 
  hc_add_theme(hc_theme_google()) %>%
  hc_title(text = "Summary statistics by Day of pickup :)") %>%
  hc_plotOptions(
    series = list(
      showInLegend = FALSE, 
      pointFormat = "{point.y}%"
    ), 
    column = list(
      colorByPoint = TRUE
    )
  ) %>% 
  hc_yAxis(title = list("Values/day"), 
           labels = list(format = "{value}"))   %>%
  hc_subtitle(text = "Summary statistics grouped by day") %>%
  hc_credits(
    enabled = TRUE, text = "Damiano ;p/click",
    href = "https://www.kaggle.com/damianpanek"
  )


hc_by_day
```




```{r echo = FALSE}

freq_by_month_trip <- train_joined %>%
    select(pickup_month, trip_duration) %>%
    group_by(pickup_month)  %>%
    summarize(count  = n(), 
              mean_trip = mean(trip_duration, na.rm = TRUE), 
              median_trip = median(trip_duration, na.rm = TRUE), 
              sd_trip = sd(trip_duration, na.rm = TRUE))



datatable(freq_by_month_trip)
```


```{r echo = FALSE}
hc_by_month <- highchart() %>%
  hc_plotOptions(
    series = list(
      showInLegend = FALSE, 
      pointFormat = "{point.y}%"
    ), 
    column = list(
      colorByPoint = TRUE
    )
  ) %>% 
  highchart() %>%
  hc_add_series(freq_by_month_trip, "line",  hcaes(x = pickup_month, y = mean_trip),name = "Mean") %>%
  hc_add_series(freq_by_month_trip,   "line" , hcaes(x=  pickup_month,  y= median_trip), name = "median") %>%
  hc_add_series(freq_by_month_trip, "line", hcaes(x =  pickup_month, y = sd_trip), name = "sd") %>% 
  hc_xAxis(categories = c("2016-01", "2016-02", "2016-03", "2016-04", "2016-05", "2016-06")) %>%
  hc_add_theme(hc_theme_google()) %>%
  hc_title(text = "Summary statistics by Month of pickup :)")
  
hc_by_month                


```

# Leaflet section

First we have to rbind drop and pick rows to create sequence. 
Then i decide to create leaflet using makecluster options. 
In my opinion it looks very nice ^^ last leaflet plot is a try and it will be updated, but id you are bored 
then you can draw a line and check it actual length ;p


Update : drawing in leaflet does not work  on kaggle :(


```{r echo = FALSE}

library(leaflet)
library(leaflet.extras)


lon_lat <- train_joined[, c("pickup_longitude", "pickup_latitude", 
                            "dropoff_longitude", "dropoff_latitude")]





lon_lat$rown <- as.numeric(rownames(lon_lat))



lon_min <- lon_lat[rown < 300 ,]

str(lon_min)
drop <- lon_min[, c("pickup_longitude", "pickup_latitude", "rown")]
pick <- lon_min[, c("dropoff_longitude", "dropoff_latitude", "rown")]


colnames(drop)  <- c("lon", "lat", "rown")
colnames(pick) <- colnames(drop)


all_bin_min <- bind_rows(drop, pick)


all_bin_min$rown2 <- rep(1:nrow(all_bin_min)+1/2,each = 2)





leaflet(data = all_bin_min) %>% addTiles() %>%
  addCircles(~lon, ~lat) %>%
  addPolygons(data = all_bin_min, lng = ~lon, 
               lat = ~lat, 
               stroke = 0.03, color =  "blue", weight = 0.4, 
               opacity = 1.2)  %>% enableMeasurePath() 





```


Leaflex plot with makecluster options 


```{r echo = FALSE}
 leaflet(data = train_joined[1:50000, ]) %>% addTiles() %>%
  addMarkers(~pickup_longitude, ~pickup_latitude, clusterOptions = markerClusterOptions()) 
```



Leaflet heatmap 


```{r echo = FALSE}

train_count <- train_joined %>% 
                select(pickup_latitude, pickup_longitude) %>%
                group_by(pickup_latitude, pickup_longitude) %>%
                summarize(count = n())


train_count <- train_count[train_count$count >1,]



 leaflet(data = train_count) %>% addTiles() %>% 
 addHeatmap(lng = ~pickup_longitude, lat = ~pickup_latitude, intensity = ~count,
             blur = 20, max = 0.05, radius = 15)


```



# Pickup grouped by month

```{r echo = FALSE}

train_count <- train_joined %>% 
                select(pickup_latitude, pickup_longitude, pickup_month) %>%
                group_by(pickup_latitude, pickup_longitude, pickup_month) %>%
                summarize(count = n())


train_count <- train_count[train_count$count >1,]



 leaflet(data = train_count) %>% addTiles() %>% 
 addHeatmap(lng = ~pickup_longitude, lat = ~pickup_latitude,
 layerId = ~pickup_month, group = ~pickup_month, intensity = ~count,
             blur = 20, max = 0.05, radius = 15)

```



    
Frequency by  day of week :)

```{r}

count_weekday <- train_joined %>%
                  select(weekday) %>%
                  group_by(weekday) %>%
                  summarize(count = n())

count_weekday <- data.table(count_weekday)


count_weekday <- count_weekday[is.na(weekday)  ==  FALSE, ]

count_weekday <- data.frame(count_weekday)

tm <- treemap(count_weekday , index = c("weekday"),
              vSize = "count")



hctreemap(tm)

```

```{r}


tm

```
todo : complex maps
todo: new dataset

